---
title: "Google Case Study: Wellness Technology Company"
author: "Idowu Efunogbon"
date: "August 11, 2022"
output: rmarkdown::github_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Bellabeat: How can a wellness technology company play it smart?



In this case study, I'll demonstrate the steps to solve the Bellabeat Wellness company  case study using  Excel Spreadsheet, and R.

The following six phase of data analysis process is used for the case study:  Ask, Prepare, Process, Analyze, Share, and Act.




#### **Table of Content:**

Summary

Characters and Products

Characters

Bellabeat Products

Phase 1: Ask 

Business Task

Phase 2: Prepare 

Data Availability

Phase 3: Process

Phase 4: Analyze

Phase 5: Share

Phase 6: Act  (Conclusions and Recommendation)

#### **SUMMARY**
I am  a Junior Data Analyst working on the Marketing Analyst Team at Bellabeat. Bellabeat which was founded in 2013, is a high-tech company that manufactures health-focused smart products for women around the world. These smart products collect women data on activity, sleep, stress and reproduction health and empower them with knowledge about their own health and habits. 
```{r}

```

#### **CHARACTERS AND PRODUCTS**

**CHARACTERS**

**Urška Sršen:** Bellabeat’s Cofounder and Chief Creative Officer

**Sando Mur:** Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team

**Bellabeat marketing analytics team**





**BELLABEAT PRODUCTS**

**Bellabeat app**: The Bellabeat app provides users with health data related to their activity, sleep, stress, menstrual cycle, and mindfulness habits. This data can help users better understand their current habits and make healthy decisions. The Bellabeat app connects to their line of smart wellness products.

**Leaf:** Bellabeat’s classic wellness tracker can be worn as a bracelet, necklace, or clip. The Leaf tracker connects to the Bellabeat app to track activity, sleep, and stress.

**Time:** This wellness watch combines the timeless look of a classic timepiece with smart technology to track user activity, sleep, and stress. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness.

**Spring:** This is a water bottle that tracks daily water intake using smart technology to ensure that you are appropriately hydrated throughout the day. The Spring bottle connects to the Bellabeat app to track your hydration levels.

**Bellabeat membership:** Bellabeat offers a subscription-based membership program for users. Membership gives users 24/7 access to fully personalized guidance on nutrition, activity, sleep, health and beauty, and mindfulness based on their lifestyle and goals.


#### **PHASE 1: ASK:**

Why conduct the analysis, what problem am I trying to solve, and defined the problem by understanding the stakeholders expectation.


**Why conduct the analysis:**
- What are some trends in smart device usage?
- How could these trends apply to Bellabeat customers?
- How could these trends help influence Bellabeat marketing strategy?


**Business Task:**
Identify trends on how consumers use their non-bellabeat smart device, and make recommendation on how Bellabeat marketing team can leverage on my analysis to improve the bellabeat  marketing strategy.


**Stakeholders:**
Urška Sršen: Bellabeat’s Cofounder and Chief Creative Officer

Sando Mur: Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team

Bellabeat marketing analytics team




#### **PHASE 2: PREPARE:**

**Data:** The Data used for carrying out this analysis is the [Fitbit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit) (CC0: Public Domain, dataset made available through [Mobius](https://www.kaggle.com/arashnic)).It is a public dataset and can be downloaded from Kaggle website.

**Data Organization:** The Fitbit Dataset; The folder named Fitabase Data 4.12.16-5.12.16 is a structured dataset that contains 18 CSV files. Each csv file have different quantitative data on women physical activity, heart rate, and sleep monitoring. The dataset ( in each CSV file) is in long format that is it contains values that do repeat in the first column. The value in the first column which is the user ID are repeated.


**Data Credibility/Limitation:** Users demographics in the dataset is unknown. The total number of Fitbit users that provided their data is 30.<https://zenodo.org/record/53894#.YvMtAPjMK3C>. Therefore, we are dealing with sampling bias because the data do not contains the adequate information needed. Another limitatation is that the dataset was generated in year 2016 and it took 2 months  to carry out the survey. The dataset which was made available for this case study is available through Mobius on kaggle website which is termed third - party data. 

**Data licensing & privacy:** The data is an open-source data (license is CCO: public Domain) from Kaggle website that is the data is  available to the public to use for analysis.


#### **PHASE 3: PROCESS:**

I will be using R and SQL to carry out my analysis. visualization and report.

After exploring the 18 csv files, I streamline my analysis to using 4 csv files. The following are the csv files  for my case study:

*dailyActivity_merged*

*Steps_merged*

*sleepDay_merged*

*weightLogInfo_merged*

 **Installing packages and opening libraries**
The following packages  are loaded in order to carry out my analysis:

```{r}
# opinionated collection of R packages designed for data science
library(tidyverse)


library(lubridate)

 #for data visualization
library(ggplot2)    

# a structure of data manipulation that help resolve the most frequent data
library("dplyr")

```


**Importing Datasets**



```{r}

#For this project I will be loading 4 csv files out of the 19 sheets for my analysis

daily_activity <- read.csv("C:/Desktop/Analytics Certificate/Course 8/dailyActivity_merged.csv")
daily_steps <- read.csv("C:/Desktop/Analytics Certificate/Course 8/dailySteps_merged.csv")
weight_log_info <- read.csv("C:/Desktop/Analytics Certificate/Course 8/weightLogInfo_merged.csv")
daily_sleep <- read.csv("C:/Desktop/Analytics Certificate/Course 8/sleepDay_merged.csv")


# To display the structure of each columns in daily activity 

str(daily_activity)

# To display the columns and first serveral rows in daily activity datset.
head(daily_activity)

# To display the column name of daily_activity dataframe
colnames(daily_activity)

# columns of the dataset and display some portion of the data with respect to each attribute
library(tidyverse)

glimpse(daily_activity)



# To display the summary of each columns in daily steps
str(daily_steps)

# To display the columns and first serveral rows in daily daily steps.
head(daily_steps)

# To display the column name 
colnames(daily_steps)

# columns of the dataset and display some portion of the data with respect to each attribute
glimpse(daily_steps)


# To display the structure of each columns in daily weight log info
str(weight_log_info)

# To display the columns and first serveral rows in daily weight log info.
head(weight_log_info)

# To display the column name 
colnames(weight_log_info)

# columns of the dataset and display some portion of the data with respect to each attribute
glimpse(weight_log_info)

# To display the structure of each columns in daily sleep
str(daily_sleep)

# To display the columns and first serveral rows in daily daily sleep.
head(daily_sleep)

# To display the column name 
colnames(daily_sleep)

# columns of the dataset and display some portion of the data with respect to each attribute
glimpse(daily_sleep)

# Cleaning and Formatting

# To check for error and inconsistency in the 3 tables that will be used for thea analysis
length(unique(daily_activity$Id))   # for daily_activity


length(unique(daily_sleep$Id))     # for daily_sleep

length(unique(weight_log_info$Id))    # for weight_log_info

length(unique(daily_steps$Id))     # For daily Steps

```



#### To check for duplicates in daily activity

```{r}
sum(duplicated(daily_activity))
```
No duplicate value



#### To check for duplicates in daily steps

```{r}
sum(duplicated(daily_steps))
```
No duplicate value





#### To check for duplicate in daily sleep

```{r}
sum(duplicated(daily_sleep))
```
Found 3 duplicate values in daily_sleep table.







#### To remove the 3 duplicate  vslues found in daily sleep and rename the dataframe

```{r}
new_daily_sleep <-daily_sleep[!(duplicated(daily_sleep) | duplicated(daily_sleep, fromLast = TRUE)), ]
sum(duplicated(new_daily_sleep))
```
 No duplicate value
 
 
 The "duplicated" function is used to  test to if a line appears at least for the second time starting at line one. 
Both boolean results are combined with | (logical 'or') into a new vector which indicates all lines appearing more than once. The result of this is negated using ! thereby creating a boolean vector indicating lines appearing only once.



#### To check for duplicate in weight log info

```{r}
sum(duplicated(weight_log_info))
```
No duplicate value



#### To check for missing values in 4 csv files

```{r}
any(is.na(daily_activity)) # check for missing values in daily activity dataframe

any(is.na(new_daily_sleep)) # check for missing values in new daily sleep dataframe

any(is.na(daily_steps)) # check for missing values in daily steps dataframe

any(is.na(weight_log_info)) # check for missing values in weight log info dataframe
```

#### To remove NA column in weight log info 

```{r}
library("dplyr")   

new_weight_log_info <- select(weight_log_info, -Fat) # The "-" tell the select function that the column fat should be drop from the dataframe
head(new_weight_log_info)
```

#### To check for missing values in weight log info dataframe

```{r}
any(is.na(new_weight_log_info)) 
```
 No more missing value that is "False"
 
 
I noticed the date column in the 4 dataframe are store in character form, so it need to be converted to datetime type.Also, the 4 datasets for this analysis  have the same column named "Id"


#### To check the number of distinct Id in daily_activity, daily_sleep and weight _log_info data frames
```{r}

n_distinct(daily_activity$Id)

n_distinct(daily_sleep$Id)

n_distinct(new_weight_log_info$Id)

n_distinct(daily_steps$Id)
```
daily_activity and daily_steps have the same number of observations




#### To check the number of observation in daily_activity, daily_sleep and weight _log_info data frames
```{r}

nrow(daily_activity)

nrow(daily_sleep)

nrow(new_weight_log_info)

nrow(daily_steps)
```
There are 940 records in daily_activity data, 413 in daily_sleep, 67 in weight_log_info and 940 records in daily_steps



Since the date column in the 4 dataframe are store in character form, it need to be converted to a datetime type.
To create month, date and weekday columns  into activity_date, weight_log_info and daily_sleep columns





#### Convert the char ActivityDate column to Date format
```{r}
# daily activity 
daily_activity$Date <- as.Date(daily_activity$ActivityDate,"%m/%d/%Y")
daily_activity$Date
```

#####  To create the month column using the conversion date dataframe
```{r}

daily_activity$Month <- format(daily_activity$Date,"%B")
daily_activity$Month

```

####  To create the weekday  column using the conversation date dataframe
```{r}
daily_activity$WeekDay <- format(daily_activity$Date,"%A")
daily_activity$WeekDay

```

####  Convert the char SleepDay column to Date format
```{r}
# new daily sleep 
new_daily_sleep$Date <- as.Date(new_daily_sleep$SleepDay,"%m/%d/%Y")
new_daily_sleep$Date

```

#### To create the month column using the conversion date dataframe
```{r}
new_daily_sleep$Month <- format(new_daily_sleep$Date,"%B")
new_daily_sleep$Month

```



#### To create the weekday  column using the conversation date dataframe
```{r}
new_daily_sleep$Weekday <- format(new_daily_sleep$Date,"%A")
new_daily_sleep$Weekday 

```



#### Convert the char Date column to Date format
```{r}
# weight_log_info
new_weight_log_info$New_Date <- as.Date(weight_log_info$Date,"%m/%d/%Y")
new_weight_log_info$New_Date

```



#### To create the month column using the conversion date dataframe
```{r}
new_weight_log_info$Month<- format(new_weight_log_info$New_Date,"%B")
new_weight_log_info$Month

```


#### To create the weekday  column using the conversation date dataframe
```{r}
new_weight_log_info$WeekDay <- format(new_weight_log_info$New_Date,"%A")
new_weight_log_info$WeekDay

```



#### **PHASE 4 AND  PHASE 5: ANALYZE AND SHARE**

#### 1. DAILY ACTIVITY

#### Summary of daily_activity

```{r}
summary(daily_activity)


```

The average mean count of daily_activity daily is 7638 and a total distance of 5.490km
 Average mean  of active minutes daily is 21.16;
Average mean of Sedentary minuts daily is 991.2'
The average calories consumed daily is 2304 kCal
The average totalsteps is 7638



Since their is no demographic variables to classify how many steps is considered to be active in order to determine the type of user we have. I will be using the the article titled “How many steps a day is considered active”
by Medical Author: Pallavi Suyog Uttekar, MD and reviewed on 7/15/2021 by Shaziya Allarakha, MD (https://www.medicinenet.com/how_many_steps_a_day_is_considered_active/article.htm).to classify the users.

Sedentary: Less than 5,000 steps per day

Low active: about 5,000 to 7,499 steps daily

Somewhat active: about 7,500 to 9,999 steps daily

Active: More than 10,000 steps daily

Highly active: More than 12,500 steps daily



#### To create a new table called usertype, we first calculate the daily_steps_average per user

```{r}
daily_step_average <- daily_activity %>% 
  group_by(Id) %>% 
  summarise (mean_daily_steps = mean(TotalSteps))


head(daily_step_average)

```


#### create the new table called usertype and classify the user based on the daily step average

```{r}
User_Type  <-daily_step_average %>%
  mutate(User_Type = case_when(
    mean_daily_steps< 5000 ~"Sedentary",mean_daily_steps >= 5000 & mean_daily_steps < 7499 ~ "Low Active",
    mean_daily_steps >= 7500 & mean_daily_steps < 9999 ~ "Somewhat Active",  mean_daily_steps >= 10000 ~ "Active",
    mean_daily_steps > 12500 ~ "Highly Active"))

head(User_Type)

```



#### create the percentage of each user type 

```{r}
User_Type_percent<- User_Type %>%
  group_by(User_Type) %>%
  summarise(Total= n()) %>%
  mutate(labels =scales::percent(Total / sum(Total))) %>% 
  arrange(desc(labels))

head(User_Type_percent)

```

#### Plot the graph of each user type

You can also embed plots, for example:

```{r}
ggplot (data= User_Type_percent, aes(x=User_Type, y=labels))+
  geom_col()+labs(x="User Type",y="Total %",title="User Type Distribution by Mean of Daily Steps")
```


From the graph, Low Active (27.3%) and Somewhat Active(27.3%)  are relatively distributed among the user. 




#### 2. WEIGHT LOG INFO

```{r}

#Summary of weight_log_info
summary(new_weight_log_info)

```

The average weightkg is 72.04kg
Average BMI is 25.19


#### 3. DAILY SLEEP
```{r}

#Summary of daily_sleep
summary(new_daily_sleep)

```
The mean total minutes  asleep is 418.9
The average total time in bed is 458.3 minutes


##### 4. DAILY STEPS
```{r}

#Summary of daily_steps
summary(daily_steps)

```
The average steptotal is 7638 steps



#### To calculate the most active days of the week 

#### using the totalsteps by weekday
```{r}

# 1. Order the day of the week
daily_activity$WeekDay <- ordered(daily_activity$WeekDay,levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))

ggplot(data = daily_activity, aes(x= WeekDay, y = TotalSteps))+
  geom_col(fill="purple") + labs(x="WeekDay",y="TotalSteps",title = "Week Day by Total Steps")

```


User use FitBit fitness tracker application to track thier totalsteps  mostly Tuesday, Wednesday, Thursday.This behaviour can be term as a result of users using  the tracker to tack to track thierq daily activities.



#### using the VeryActiveMinutes by weekday
```{r}

#2.  Order the day of the week
daily_activity$WeekDay <- ordered(daily_activity$WeekDay,levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))


ggplot(data = daily_activity, aes(x= WeekDay, y = VeryActiveMinutes))+
  geom_col(fill="blue") + labs(x="WeekDay",y="VeryActiveMinutes",title = "Week Day by Very Active Minutes")


```



#### using the Calories by weekday
```{r}
# 3. Order the day of the week
daily_activity$WeekDay <- ordered(daily_activity$WeekDay,levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))


ggplot(data = daily_activity)+ geom_col(aes(x= WeekDay, y = Calories),fill= "orange")+
  labs(x="WeekDay",y="Calories",title = "Week Day by Calories")
```


User use FitBit fitness tracker application to track thier calories and Tuesday, Wednesday , Thursday have the hightest caloires burned.




#### 1. for Daily activity (Relation between Total Steps and SedentaryMinutes)
```{r}
ggplot (data= daily_activity, aes(x=TotalSteps, y=SedentaryMinutes,color= Calories))+
  geom_point()+ labs(x="Total Steps",y="Sedentary Minutes",title="Total Steps by Sedentary Minutes")

```


From the scatter graph, there is a negative relationship between total steps and sedentary miutes because people do not move when they are not physically active


#### 2. for Daily activity (To find if there is a relation between Total Steps by calories and colour of the line can be controlled using colour aesthetic
```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y = Calories))+ geom_point() + stat_smooth(method=lm, fill="blue", colour="darkblue", size=1)+
  labs(x="Total Steps",y="Calories",title="Total Steps by Calories")


```


From the plotted scatter graph, there is a positivecorrelation between the totalsteps and calories people that took the most total steps burn the most calories i.e the higher the steps the more calories user burn. 


#### 3.  To find if there is a relation between sleep and time in bed?
```{r}
ggplot(data= new_daily_sleep, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + geom_point() + stat_smooth(method = lm,  fill="blue", colour="darkblue", size=1)+
  labs(x="Total Minutes Asleep",y="Total Time In Bed",title="Total Minutes Asleep by Total Time In Bed")

```


From the scatter graph, we noticed a strong positive correlation between TotalMinutesAsleep and TotalTimeInBed also  we observed some outliers in  the middle and top of the scatter plot.
we can say the outliers are the people who spend lot of time in bed but did not actually sleep  due to different reasons. To improve the sleep hours, a beep on the bellet time watch should be use to prepare user to sleep.



#### 4.  To find if there is Relationship between  active minutes  and calories

```{r}
ggplot(data=daily_activity,aes(x = VeryActiveMinutes, y = Calories, color = Calories)) + geom_point() + stat_smooth(method = lm,  fill="orange", colour="darkorange", size=1)+
labs(x="Very Active Minutes",y="Calories",title = "Very Active Minutes by Calories Burned")
```


From the graph, there is a strong correlation between very active minutes and calories burned with some outliers at bottom left and top left of the plot.




#### To plot daily sleep per week day

```{r}
# Order the day of the week
new_daily_sleep$Weekday <- ordered(new_daily_sleep$Weekday,levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))


ggplot(data = new_daily_sleep, aes(x= Weekday, y = TotalMinutesAsleep))+
  geom_col() + labs(x="Weekday",y="Total Minutes Asleep",title = "Total Minutes Asleep By Week Day")
```


User sleep most on Wednesday, and Tuesday  while the least sleep day is on Monday.





#### To analyze the relationship of weight data, merge dailyactivity and new weight log info dataframes

```{r}
weight_daily_activity <- merge(daily_activity,new_weight_log_info, by = "Id")
head(weight_daily_activity)

summary(weight_daily_activity)
```





####  To plot the relationship between veryactiveminutes and weightin kg

```{r}
ggplot(data=weight_daily_activity) + geom_violin(mapping=aes(x=VeryActiveMinutes, y=WeightKg), fill = "blue") +
labs(title="Violin plot", subtitle="User Weight by Very Active Minutes", x="Very Active Minutes", y="Weight in Kg")


```


From the violin plot, the median weight is 63.

Most users have a weight that is between 61 and 85, but some users have weight that are low as 53 and as high as 134. From the plot we can see outlier weight to be 134.



#### **PHASE 6: ACT (CONCLUSION AND RECOMMENDATIONS)**

From the analysis carried out and using the business questions:


- The average total steps per day is 7638 which is below the 10,000 recommended steps per day which is termed the gold standards for adults. Therefore, the app should track users who do not meet the 10,000 recommended steps per day and send a reminder to to encourage them  

- People tend to be more active on on Tuesday, Wednesday and Thursday so they tend to use their app on week days maybe because they spend more time outside compare to weekends

- Most users weight in Kg is between 61 - 85. Since bellabeat focus is on young and adult women empowerment and the data lack information on women Gender and Age therefore, there is need for more data on young and adult female gender, and age  in order to meet  bellabeat mission. 

- Bellabeat marketing team can encourage users by educating them  the benefits of the fitness, recommend the types of exercises users can partake in, calories intake and burn rate information on bellabeat application.

- The Bellebeat application  should always send reminder during weekends to encourage users to exercise regularly.

- The Bellabeat app should create a reward system such that users who meet their daily steps goals would earn points which can be converted to discount for bellabeat membership  at the end of every year 

